---
title: "STAT6730 Project"
subtitle: "Exploring park access data from the Trust for Public Land"
author: "Polina Berezina, Shan Tang"
date: "12/14/2021"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, fig.align = 'center', dpi=300) # global settings
options(htmltools.dir.version = FALSE) 
# install.packages(c("ggmap", "tidygeocoder", maps", "mapdata", "VIM", "reshape2", "mice"))
library(dplyr)
library(ggplot2)
library(tidyverse)
library(mice)
library(reshape2)
library(VIM)
library(purrr)
library(xaringan)
library(ggmap)
library(tidygeocoder)
library(maps)
library(mapdata)
```
  
## Introduction
The Trust for Public Land (TPL) [1] tracks green space availability across U.S. using the proprietary **ParkScore index**. This index measures how well cities are meeting their residents’ need for public parks based on several categories:

- **access**: percent of residents that live within a 10-minute walking distance, or 0.5 miles, to a public park.
- **acreage**: median size, percent of parkland within the city area.
- **investment**: sum of per capita statistic on public, non-profit spendings to manage parklands, volunteer hours. 
- **amenities**: per capita statistics on the number of basketball hoops, dog parks, playgrounds, restrooms, recreation and senior centers, and splashpads.
- **equity** (2021 only): average of four metrics. Per capita ratio of nearby park space between neighborhoods of color and white, as well as low-income vs. high-income neighborhoods. Percentage of people of color living within a 10-minute walk to a park and a similar percentage for low-income households.

Each city is awarded up to 100 points based on each one of these categories, and total of the scores is normalized to form an overall ParkScore rating. The data is taken from the Tidy Tuesday project [2]. In this report, we would like conduct exploratory data analysis to explore yearly trends across different categories of ParkScore and how the TPL has changed their methodology. We also explore importance of certain park features to their rankings. Finally, we analyze the newest 2021 data published by TPL to uncover how addition of the equity category to the ParkScore has affected cities' rankings.


```{r}
raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-22/parks.csv')
```

Regional patterns 

WOULD NOT SHOW UP TO SPEED UP TESTING. REMOVE "EVAL=FALSE" IN THE END

```{r map,  fig.align='center', fig.width=10,fig.height=6, eval = FALSE}
city_locations <- select(raw, city, rank, year) %>% 
  filter(year == 2020, !city %in% c('Honolulu', 'Anchorage')) %>% # so that we show contiguous usa only, for 2020 year
  mutate(address = paste0(city,', USA'))  %>%  # without appending USA, algorithm confuses same-named cities in other countries
  geocode(address, method = 'osm', lat = lat, long = long) # geocode, or find lat long for each city

usa <- map_data("usa")
states <- map_data("state")

mybreaks <- c(1, 25, 50, 75)
ggplot() + 
  geom_polygon(data = states, aes(x = long, y = lat, group = group), color="grey", fill="grey99") +
  coord_fixed(1.3) +
  geom_point(data=city_locations, aes(x=long, y=lat,  color=rank, size=rank), stroke=F, alpha=0.8) +
  #scale_alpha_continuous(range=c(0.1, 0.9)) + 
  scale_colour_gradient2(name="Rank", low ="#91cf60", mid="#ffffbf", high = "#fc8d59", midpoint = 47, breaks=mybreaks, labels = c("1-19", "25-50", "50-74", "75+"), guide = "legend") +
  scale_size_continuous(name="Rank", range = c(6,2), breaks=mybreaks, labels = c("1-19", "25-50", "50-74", "75+"))+
  guides(fill=FALSE) +
  labs(caption = "Park access rankings in contiguous USA in 2020") +
  theme(plot.margin = unit(c(-2, 0, -2, -.1), "cm")) +
  theme_void()
```


## Data cleaning and imputation

- Year as a factor
- Removed % and $ in the data frame, convert variables from string (%) to double (0-1)
- rename the column names

```{r cleaning1}
raw$year <- as.factor(raw$year)
original.colnames <- colnames(raw) # save for the 2021 data used later

library(stringr)
# replace with dplyr?
raw$park_pct_city_data<- as.numeric(str_replace_all(raw$park_pct_city_data, "%", ""))
raw$pct_near_park_data<- as.numeric(str_replace_all(raw$pct_near_park_data, "%", ""))
raw$spend_per_resident_data<- as.numeric(gsub("\\$","",raw$spend_per_resident_data))

raw1<-raw
colnames(raw1)<-c("year","rank","city","med park size","park size score1","park size pct", "park size score2",
                 "park access","park access socre","park investment","park investment score",
                 "basketball","basketball score", "dogpark",  "dogpark score", "playground", "playground score",
                 "rec sr","rec sr socre","restroom", "restroom score" , "splashground",  "splashground score",
                 "amenities score", "total scores", "total%","city dup","park benches")
#head(raw1[,c(1:3,6,8,10)], 1)
```


- More than 30% of data is missing in 6 out of 28 variables, including restroom, splashground and park benches counts. We omit these variables from analysis. 
- Assuming missing at random for others: impute using mice package with PMM Hot Deck method

```{r remove,fig.width=10,fig.height=5}
aggr(raw1[,12:28], cex.axis=0.65)
raw1 <- select(raw1, -c("city dup", "park benches", "splashground", "splashground score", "restroom score", "restroom"))

raw2 <- select(raw, -c(city, rank)) # everything but city and rank - not useful in imputations
imp <- mice(raw2, method="pmm", donors=5, print=FALSE) 
df <- complete(imp, 1) %>% mutate(., city = raw$city, rank = raw$rank, .after=1) # append back city and rank vars
```


## Data analysis

Distribution of variables

```{r explore1, fig.width=10,fig.height=6}
#basic histogram
df1<-df
colnames(df1)<-c("year","city","rank","med park size","park size score1","park size pct", "park size score2",
                 "park access","park access socre","park investment","park investment score",
                 "basketball","basketball score", "dogpark",  "dogpark score", "playground", "playground score",
                 "rec sr","rec sr score","restroom", "restroom score" , "splashground",  "splashground score",
                 "amenities score", "total scores", "total%","city dup","park benches")
melt(df1) %>% 
  ggplot(.,aes(x = value)) +  ##,mapping = aes(value, after_stat(density))
      facet_wrap(~variable,scales = "free_x") + 
      geom_histogram() +    #geom_histogram() + # geom_boxplot()  stat_bin(binwidth = h_sturges)
    theme_classic() +
    theme(strip.text.x = element_text(size = 11), text = element_text(size = 9), axis.text.x = element_text(angle=90)) 
```


Variables over the years. Increasing trend in `park size`, `park access`, `playground` scores

```{r explore2, fig.width=10,fig.height=6}
# boxplots per year for every varaible
melt(df1) %>% 
  ggplot(.) + 
    facet_wrap(~variable, scales = "free_y") + 
    geom_boxplot(aes(y=value, x=year, group = year)) +
  theme_classic() +
  theme(strip.text.x = element_text(size = 11),text = element_text(size = 9), axis.text.x = element_text(angle=90)) 
```

Park Scores vs. Features. TPL has different max. scores or feature weight across years. <span style="color:purple">2020-21</span> scores are higher across park size, easier accessibility and investment in parks features.

```{r, fig.width=10,fig.height=6}
library(ggpubr)
##use function to reduce copy and past
plot_trend<-function(df_sub, labsname){
  ##labsname = c(title, xlab, ylab)
  pp<- ggplot( df_sub, aes_string(x=names(df_sub)[1], y=names(df_sub)[2], color=names(df_sub)[3]) )+ 
    geom_point(alpha=0.6)+
    labs(title=labsname[1], x=labsname[2], y=labsname[3])+
    theme_classic() +
    theme(text = element_text(size = 9))
    
  return(pp)
}

p1<- df %>% 
  select(med_park_size_data,med_park_size_points,year) %>% 
  plot_trend (labsname=c("Size Score 1 VS. Median Park Size", "Median Park Size (Acres)", "Score") )

p2<-df %>% 
  select(park_pct_city_data,park_pct_city_points,year) %>% 
  plot_trend (labsname=c("Size Score 2 VS. % Park Size", "Park Size (% city area) ", "Score") ) 
  
p3<- df %>% 
  select(pct_near_park_data,pct_near_park_points,year) %>% 
  plot_trend (labsname=c("Access Score VS. Accessibility", "Residents% within 10-min Walking", "Score") )
  
p4<- df %>% 
  select(spend_per_resident_data,spend_per_resident_points,year) %>% 
  plot_trend (labsname=c("Cost Score VS. Investment", "Cost per resident ($)", "Score") )

ggarrange(p1, p2, p3, p4,
          ncol = 2, nrow = 2,
          common.legend = TRUE, legend="bottom")
```


Amenities: (number of) basketball hoops, dog parks, playgrounds. Overall increasing trend in score with better amenities. But plateaus

```{r}
p1<- df %>% 
  select(basketball_data,basketball_points,year) %>% 
  plot_trend (labsname=c("Basketball Hoops", " ", "Score") )
```


```{r, fig.width=10,fig.height=4}
p1<- df %>% 
  select(basketball_data,basketball_points,year) %>% 
  plot_trend (labsname=c("Basketball Hoops", " ", "Score") )

p2<-df %>% 
  select(dogpark_data,dogpark_points,year) %>% 
  plot_trend (labsname=c("Dog Parks", " ", "Score") ) 
  
p3<- df %>% 
  select(playground_data,playground_points,year) %>% 
  plot_trend (labsname=c("Playgrounds", " ", "Score") )

ggarrange(p1, p2, p3, 
          ncol = 3, nrow = 1,
          common.legend = TRUE, legend="bottom")
```


#### Top ranking cities across years

```{r, fig.width=10,fig.height=6}
city_name<-union(df$city[df$rank<=10 & df$year==2015], df$city[df$rank<=10 & df$year==2020])  ## city
city_name<-city_name[city_name %in% c("Arlington, Virginia","Minneapolis","Irvine") ==FALSE]
ggplot(subset(df, city %in% city_name & year %in% c(2015, 2020)),
       aes(x=rank, y=city)) +
    geom_point(aes(color=year),size=3.5)+
    geom_hline(yintercept=c(1:12), linetype="dashed")+
    labs(x="Park rank", y="City")+
    theme_classic() +
    theme(text = element_text(size = 13))
```



#### Exploring equity in 2021

TPL added 4 features that go into the new **equity score** : park space ratio and 10-minute walk for low-income and people of color.


Hard to directly compare statistics since they changed weighting methods.

```{r equity}
df.2021 <- data.frame(read.csv('parks_2021_clean.csv'))
df.2021$city <-str_sub(df.2021$city, 1 ,nchar(df.2021$city)-4) #remove state
df.2021$year <- rep(2021, nrow(df.2021)) # create a year variable for plotting purposes

df.2020 <- filter(df, year == 2020)
cbind(seq(1,10), df.2020[1:10,2], df.2021[1:10,2])
```



## Conclusion 

Pay attention to the different scores range/weight across the years

Parks tend to have greater score with larger park size, easier accessibility, higher investment and more facilities. 

TPL affected rankings of a lot of cities by including an equity metric in 

## References

[1] https://www.tpl.org/parkscore

[2] https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-06-22/readme.md
