---
title: "STAT6730 Project"
subtitle: "Exploring park access data from the Trust for Public Land"
author: "Polina Berezina, Shan Tang"
date: "12/13/2021"
output: 
  html_document:
    number_sections: true

---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, dpi=300) # setting global settings
options(htmltools.dir.version = FALSE) 
# install.packages(c("ggmap", "tidygeocoder", maps", "mapdata", "VIM", "reshape2", "mice"))
library(dplyr)
library(ggplot2)
library(tidyverse)
library(mice)
library(reshape2)
library(VIM)
library(purrr)
# for mapping 
library(ggmap)
library(tidygeocoder)
library(maps)
library(mapdata)
```
  
# Introduction
The Trust for Public Land (TPL) tracks green space availability across U.S. using the **ParkScore index** to measures how well cities are meeting their residents’ need for public parks based on: 
- **park access**: residents% walk to park within 10-min walking 
- **acreage**: median size, parkland% of city area
- **investment**: spending/resident
- **amenities**: basketball hoops, dig park, playground, restroom, etc.
- **equity** distribution of parks according to race and income (2021 only)

### We explore
- Yearly trends
- The relationship between some features and park rankings/scores
- Affect of adjusting for equity in the newest ranking in 2021

# Ranking of cities in 2020

Regional patterns?

```{r map, fig.align='center', fig.width=10,fig.height=6}
raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-22/parks.csv')
city_locations <- select(raw, city, rank, year) %>% 
  filter(year == 2020, !city %in% c('Honolulu', 'Anchorage')) %>% # so that we show contiguous usa only, for 2020 year
  mutate(address = paste0(city,', USA'))  %>%  # without appending USA, algorithm confuses same-named cities in other countries
  geocode(address, method = 'osm', lat = lat, long = long) # geocode, or find lat long for each city
usa <- map_data("usa")
states <- map_data("state")
ggplot() + 
  geom_polygon(data = states, aes(x = long, y = lat, group = group), color="grey", fill="grey99") +
  coord_fixed(1.3) +
  guides(fill=FALSE) +
  geom_point(data=city_locations, aes(x=long, y=lat), size = 1.4,colour = "black") + 
  geom_point(data=city_locations, aes(x=long, y=lat,  color=rank), size=1) +
  scale_colour_gradient2(low ="#fc8d59", mid="#ffffbf", high = "#91cf60", midpoint = 47) +
  theme(plot.margin = unit(c(-2, 0, -2, -.1), "cm")) +
  theme_void()
```


# Data cleaning
- Year as a factor
- Removed % and $ in the data frame, convert variables from string (%) to double (0-1)
- rename the column names

#### Before
```{r cleaning1}
raw$year <- as.factor(raw$year)
original.colnames <- colnames(raw) # save for the 2021 data used later
head(raw[,c(1:3,6,8,10)], 1)   ### show the changes that we made
```

#### After
```{r cleaning2}
library(stringr)
# replace with dplyr?
raw$park_pct_city_data<- as.numeric(str_replace_all(raw$park_pct_city_data, "%", ""))
raw$pct_near_park_data<- as.numeric(str_replace_all(raw$pct_near_park_data, "%", ""))
raw$spend_per_resident_data<- as.numeric(gsub("\\$","",raw$spend_per_resident_data))
raw1<-raw
colnames(raw1)<-c("year","rank","city","med park size","park size score1","park size pct", "park size score2",
                 "park access","park access socre","park investment","park investment score",
                 "basketball","basketball score", "dogpark",  "dogpark score", "playground", "playground score",
                 "rec sr","rec sr socre","restroom", "restroom score" , "splashground",  "splashground score",
                 "amenities score", "total scores", "total%","city dup","park benches")
head(raw1[,c(1:3,6,8,10)], 1)
```

## Handling missing data

- More than 30% of data is missing in 6 out of 28 variables, including restroom, splashground and park benches counts. We omit these variables from analysis. 
- Assuming missing at random for others: impute using mice package with PMM Hot Deck method

```{r remove,fig.width=11,fig.height=5}
aggr(raw1[,12:28], cex.axis=0.6)
raw1 <- select(raw1, -c("city dup", "park benches", "splashground", "splashground score", "restroom score", "restroom"))
```

```{r impute}
raw2 <- select(raw, -c(city, rank)) # everything but city and rank - not useful in imputations
imp <- mice(raw2, method="pmm", donors=5, print=FALSE) 
df <- complete(imp, 1) %>% mutate(., city = raw$city, rank = raw$rank, .after=1) # append back city and rank vars
```

---
# Exploratory data analysis

## The distribution of variables

```{r explore1, fig.width=10,fig.height=6}
#basic histogram
df1<-df
colnames(df1)<-c("year","city","rank","med park size","park size score1","park size pct", "park size score2",
                 "park access","park access socre","park investment","park investment score",
                 "basketball","basketball score", "dogpark",  "dogpark score", "playground", "playground score",
                 "rec sr","rec sr socre","restroom", "restroom score" , "splashground",  "splashground score",
                 "amenities score", "total scores", "total%","city dup","park benches")
melt(df1) %>% 
  ggplot(.,aes(x = value)) +  ##,mapping = aes(value, after_stat(density))
      facet_wrap(~variable,scales = "free_x") + 
      geom_histogram() +    #geom_histogram() + # geom_boxplot()  stat_bin(binwidth = h_sturges)
    theme_classic() +
    theme(strip.text.x = element_text(size = 11), text = element_text(size = 9), axis.text.x = element_text(angle=90)) 
```


## Variables over the years

```{r explore2, fig.width=10,fig.height=6}
# boxplots per year for every varaible
melt(df1) %>% 
  ggplot(.) + 
    facet_wrap(~variable, scales = "free_y") + 
    geom_boxplot(aes(y=value, x=year, group = year)) +
  theme_classic() +
  theme(strip.text.x = element_text(size = 11),text = element_text(size = 9), axis.text.x = element_text(angle=90)) 
```

We investigated the overall variable distribution and the variable distribution over the years from 2012 to 2019. For variable distributions, most variables have an approximately normal distribution with some skewness towards right. Some potential outlines were observed. In the variable distribution over years, we observed a general increasing trend in scores for all attributes. For example, the park size score has an obvious larger number for year 2019 and 2020 compared with previous year. We will explain the reason for the increase in scores in next section. 

## Park Scores vs. Features 

```{r, fig.width=10,fig.height=6}
library(ggpubr)
##use function to reduce copy and past
plot_trend<-function(df_sub, labsname){
  ##labsname = c(title, xlab, ylab)
  pp<- ggplot( df_sub, aes_string(x=names(df_sub)[1], y=names(df_sub)[2], color=names(df_sub)[3]) )+ 
    geom_point(alpha=0.6)+
    labs(title=labsname[1], x=labsname[2], y=labsname[3])+
    theme_classic() +
    theme(text = element_text(size = 9))
    
  return(pp)
}

p1<- df %>% 
  select(med_park_size_data,med_park_size_points,year) %>% 
  plot_trend (labsname=c("Median Park Size", "Median Park Size (Acres)", "Score") )
p2<-df %>% 
  select(park_pct_city_data,park_pct_city_points,year) %>% 
  plot_trend (labsname=c("% Park Size", "Park Size (% city area) ", "Score") ) 
  
p3<- df %>% 
  select(pct_near_park_data,pct_near_park_points,year) %>% 
  plot_trend (labsname=c("Accessibility", "Residents% within 10-min Walking", "Score") )
  
p4<- df %>% 
  select(spend_per_resident_data,spend_per_resident_points,year) %>% 
  plot_trend (labsname=c("Investment", "Cost per resident ($)", "Score") )

p5<- df %>% 
  select(basketball_data,basketball_points,year) %>% 
  plot_trend (labsname=c("Basketball Hoops", " ", "Score") )
p6<-df %>% 
  select(dogpark_data,dogpark_points,year) %>% 
  plot_trend (labsname=c("Dog Parks", " ", "Score") ) 
  
p7<- df %>% 
  select(playground_data,playground_points,year) %>% 
  plot_trend (labsname=c("Playgrounds", " ", "Score") )

ggarrange(p1, p2, p3, p4,p5,p6,p7,
          ncol = 4, nrow = 2,
          common.legend = TRUE, legend="bottom")
```

By plotting the park score vs. different features that include park size (median park size in acres and percentage of park size in the city), park accessibility (percentage of residents who live close to the park within 10-min walking distance), park investment (the cost for park construction per resident in dollar), and park amenities (basketball hoops, dog parks, playgrounds), an overall positive correlation between park scores and all these features were revealed. It is noticeable that different years might have different score scale for these park features. For example, the scores for park size for year 2019 and 202 range from 0 to 50. The maximum value, 50, is greater than the previous years' max values 20. We need to be aware of these changes of the score range when analyzing these data. Nevertheless, the park scores increase with larger park size, easier accessibility, greater investment and better amenities. Even though it seems obvious that good park systems need adequate acreage, access, investment and services, investigating the relationship between parks scores and specified features can be served as the guide for local park improvement efforts. 

## Top ranking cities across years

```{r, fig.width=10,fig.height=6}
city_name<-union(df$city[df$rank<=10 & df$year==2015], df$city[df$rank<=10 & df$year==2020])  ## city
city_name<-city_name[city_name %in% c("Arlington, Virginia","Minneapolis","Irvine") ==FALSE]
ggplot(subset(df, city %in% city_name & year %in% c(2015, 2020)),
       aes(x=rank, y=city)) +
    geom_point(aes(color=year),size=3.5)+
    geom_hline(yintercept=c(1:12), linetype="dashed")+
    labs(x="Park rank", y="City")+
    theme_classic() +
    theme(text = element_text(size = 13))
```

This figure shows the rank change of the top 12 parks from year 2015 to year 2020. Most cities' ranks remained unchanged from year 2015 to year 2020. San Diego had a major drop in park ranking among these cities. When we further look into different score changes for San Diego, the scores for amenities decreased from 2015 to 2020, which is a major reason that caused the rank change. 


---
## Exploring equity in 2021

- TPL added 4 features that go into the new **equity score** : park space ratio and 10-minute walk for low-income and people of color.
- Again, hard to compare statistics since they changed weighting methods. Visually:

```{r equity}
library(RCurl)
df.2021 <- data.frame(read.csv("https://raw.githubusercontent.com/pberezina/parks-project/main/parks_2021_clean.csv", header = T))
df.2021$city <-str_sub(df.2021$city, 1 ,nchar(df.2021$city)-4) #remove state
df.2021$year <- rep(2021, nrow(df.2021)) # create a year variable for plotting purposes
df.2020 <- filter(df, year == 2020)
cbind(seq(1,10), df.2020[1:10,2], df.2021[1:10,2])
```



---
## Summary & Thank you
- Pay attention to the different scores range/weight across the years
- Parks tend to have greater score with larger park size, easier accessibility, higher investment and more facilities. 
- TPL affected rankings of a lot of cities by including an equity metric in 

##### Find this project at GitHub: https://github.com/pberezina/parks-project
